---
title: "AZT1D Glycemic Control — Reproducible Analysis"
author: "Genelle Jenkins"
date: today
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
  pdf:
    toc: true
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(lubridate)

# Reads the synthetic sample data shipped with the repo (same schema as real AZT1D).
data_dir <- "../sample_data/CGM Records"
files <- list.files(data_dir, pattern = "\\.csv$", recursive = TRUE, full.names = TRUE)

cgm <- files |>
  set_names(basename(dirname(files))) |>
  map_dfr(read_csv, .id = "subject", show_col_types = FALSE) |>
  mutate(EventDateTime = ymd_hms(EventDateTime))

# ADA clinical thresholds
hypo <- 70; target_lo <- 70; target_hi <- 180; hyper <- 180
```

## Cohort overview

```{r overview}
cgm |>
  group_by(subject) |>
  summarise(
    n_readings   = n(),
    mean_glucose = mean(CGM, na.rm = TRUE),
    cv_pct       = sd(CGM, na.rm = TRUE) / mean(CGM, na.rm = TRUE) * 100,
    tir_pct      = mean(CGM >= target_lo & CGM <= target_hi) * 100,
    tbr_pct      = mean(CGM < hypo) * 100,
    .groups = "drop"
  ) |>
  arrange(desc(tir_pct)) |>
  knitr::kable(digits = 1, caption = "Per-subject glycemic summary")
```

## Time-in-range by subject

```{r tir-plot}
cgm |>
  mutate(zone = case_when(
    CGM < hypo      ~ "Below range (<70)",
    CGM > hyper     ~ "Above range (>180)",
    TRUE            ~ "In range (70-180)"
  )) |>
  count(subject, zone) |>
  group_by(subject) |>
  mutate(pct = n / sum(n) * 100) |>
  ggplot(aes(subject, pct, fill = zone)) +
  geom_col() +
  scale_fill_manual(values = c("In range (70-180)" = "#34a853",
                               "Below range (<70)" = "#ea4335",
                               "Above range (>180)" = "#fbbc04")) +
  labs(title = "Time-in-range distribution by subject",
       x = NULL, y = "% of readings", fill = NULL) +
  coord_flip() + theme_minimal()
```

## Ambulatory Glucose Profile (median + IQR by time of day)

```{r agp}
cgm |>
  mutate(tod = hour(EventDateTime) + minute(EventDateTime) / 60) |>
  group_by(tod_bin = round(tod * 2) / 2) |>
  summarise(
    med = median(CGM), lo = quantile(CGM, .25), hi = quantile(CGM, .75),
    .groups = "drop"
  ) |>
  ggplot(aes(tod_bin)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = target_lo, ymax = target_hi,
           alpha = .08, fill = "#34a853") +
  geom_ribbon(aes(ymin = lo, ymax = hi), alpha = .25, fill = "#4285f4") +
  geom_line(aes(y = med), color = "#1a73e8", linewidth = 1) +
  labs(title = "Ambulatory Glucose Profile (cohort)",
       x = "Hour of day", y = "CGM (mg/dL)") +
  theme_minimal()
```

## Takeaways

- Time-in-range and hypoglycemia exposure vary substantially across subjects.
- The AGP highlights time-of-day patterns (e.g., overnight vs. post-meal excursions).
- Reproducible in R/Tidyverse; rendered to HTML + PDF via Quarto (CI workflow included).

> Synthetic demo data. Re-point `data_dir` to DVC-pulled real data for true results.
